Path Analysis Data Development - Chapter 2

### 2016-07-15 sem-data-development.Rmd
Recalculate umca rooted within plots to use for calculating umca density.

## Examine the bay laurel data at individual and plot levels
```{r examine umca data}
bay_laurel <- filter(stems, species == "umca") %>% 
      select(plotid, cluster:stem_status, umca_slc, notes:week)
bay_laurel$species <- droplevels(bay_laurel$species)
summary(bay_laurel) # This has stem-level data on bay laurel.
length(unique(bay_laurel$plotid))# 183 plots with bay laurel
# bay_laurel <- filter(bay_laurel, plot != "BUSH01")

# Data frame of UMCA plots only
umca_plots <- bay_laurel %>%
      # select(plotid, date, species, umca_slc, year, stem_status) %>%
      group_by(plotid, year, species) %>%
      filter(species == "umca", stem_status == "Alive") %>%
      summarise(uninfected_bay_ct = length(which(umca_slc == 0)), 
                infected_bay_ct = length(which(umca_slc > 0)), 
                tot_bay = length(species)) 
summary(umca_plots)

# length(unique(umca_plots$plotid))# 183 plots with bay laurel
# umca_plots <- filter(plots_umca, plot != "BUSH01")
rm(plots_umca)
```

I can use the stem-level data to calculate plot level leaf count variables (sum, average) for each plot and year.
```{r calculate plot level leaf counts}
live_umca <- bay_laurel %>% filter(stem_status == "Alive")
filter(live_umca, is.na(umca_slc))
# Spot checking some of these shows that NAs are due to missing data, i.e. they should not be changed to zeroes
umca_plots <- left_join(
      umca_plots, 
      live_umca %>% 
            group_by(plotid, year) %>% 
            summarise(avg_lfct = mean(umca_slc, na.rm = T),
                      tot_lfct = sum(umca_slc, na.rm = T)),
      by = c("plotid","year"))
umca_plots <- left_join(
      umca_plots,
      plot_visit_dates,
      by = c("plotid","year")
)
summary(umca_plots)
```

@@

# Join the environmental data to the bay laurel infection data
umca_plots <- left_join(select(umca_plots, -species), plots_env, by = "plotid")
summary(umca_plots)
# filter(umca_plots, is.na(lon_utm))
# NAs due to no topo data for abandoned SUGAR02; remove SUGAR02
# umca_plots <- na.omit(umca_plots)
umca_plots <- rename(umca_plots, sample_year = year)
# umca_plots$plotid <- tolower(umca_plots$plotid)

# Join rainy season and summer temperatures to bay laurel infection data.
## I need to rename temperature variables to differentiate rainy vs. dry season
rs_temps <- as.tbl(rs_temps)
rs_temps$plotid <- tolower(rs_temps$plotid)
summary(rs_temps)
rs_temps <- rename(rs_temps, 
                   hrs_14_20_rs = hrs_14_20, hrs_abv25rs = hrs_abv25,
                   hrs_abv20_rs = hrs_abv20, hrs_blw10rs = hrs_blw10,
                   avg_tmax_rs = avg_tmax, avg_tmin_rs = avg_tmin)

@@

# Now I can join the data without duplicating variable names
umca_plots <- left_join(umca_plots, rs_temps, by = c("plotid", "sample_year"))
umca_plots <- left_join(umca_plots, summer_temps, by = c("plotid", "sample_year"))
summary(umca_plots)# 7 NAs introduced into the temperature variables
filter(umca_plots, is.na(avg_tmax_rs))# NAs due to records from 2003
## I don't think these will end up in my analysis, but no reason to remove them at this time.

@@

@@

umca_plots <- left_join(umca_plots, os.us.diversity, by = c("plotid"))
summary(umca_plots)
unique(filter(umca_plots, is.na(H.2014))$plotid)
## NAs due to plots abandoned/lost between 2005 and 2014; 4 unique plots
filter(umca_plots, is.na(us.H.2005))
## NAs due to plot having '0' for understory richness
filter(umca_plots, is.na(us.H.2011))
## NAs due to 2 plots being abaondoned prior to 2011 & one plot having '0' for understory richness

@@

# Add bay laurel infection data to oak plots
names(umca_plots)
oak_sod <- left_join(oak_sod, 
                     select(umca_plots, plotid, sample_year, uninfected_bay_ct:tot_lfct ), 
                     by = c("plotid", "sample_year"))
summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
oak_sod <- left_join(
      oak_sod,
      plot_visit_dates,
      by = c("plotid","sample_year"="year"))

filter(oak_sod, is.na(tot_bay), sample_year != 2003)
# Change NAs of bay laurel in oak only plots to zeroes
oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
oak_sod$infected_bay_ct[is.na(oak_sod$infected_bay_ct)] <- 0
oak_sod$tot_bay[is.na(oak_sod$tot_bay)] <- 0

@@

rs_temps
summer_temps
oak_sod <- left_join(oak_sod, rs_temps, by = c("plotid", "sample_year"))
oak_sod <- left_join(oak_sod, summer_temps, by = c("plotid", "sample_year"))
oak_sod
names(oak_sod)

>>>>

## Examine the bay laurel data at individual and plot levels
```{r examine umca data}
# All tagged bay laurel
bay_laurel <- filter(stems, species == "umca") %>% 
      select(plotid, cluster:stem_status, umca_slc, notes:week)
bay_laurel$species <- droplevels(bay_laurel$species)
summary(bay_laurel) # This has stem-level data on bay laurel.
length(unique(bay_laurel$plotid))# 183 plots with bay laurel
# bay_laurel <- filter(bay_laurel, plot != "BUSH01")

# Only bay laurel rooted within the plot boundaries
bay_laurel_in <- filter(stems, species == "umca", location == "In") %>% 
      select(plotid, cluster:stem_status, umca_slc, notes:week)
bay_laurel_in$species <- droplevels(bay_laurel_in$species)
summary(bay_laurel_in) # This has stem-level data on bay laurel.

# Data frames of UMCA-only plots, summarizing UMCA abundance and symptomatic leaf count
umca_plots <- bay_laurel %>%
      # select(plotid, date, species, umca_slc, year, stem_status) %>%
      group_by(plotid, year) %>%
      filter(stem_status == "Alive") %>%
      summarise(uninfected_bay_ct = length(which(umca_slc == 0)), 
                infected_bay_ct = length(which(umca_slc > 0)), 
                tot_bay = infected_bay_ct + uninfected_bay_ct,
                avg_lfct = mean(umca_slc, na.rm = T),
                tot_lfct = sum(umca_slc, na.rm = T)) 
summary(umca_plots)

umca_in_plots <- bay_laurel_in %>%
      # select(plotid, date, species, umca_slc, year, stem_status) %>%
      group_by(plotid, year) %>%
      filter(stem_status == "Alive") %>%
      summarise(uninfected_bay_ct = length(which(umca_slc == 0)), 
                infected_bay_ct = length(which(umca_slc > 0)), 
                tot_bay = uninfected_bay_ct + infected_bay_ct,
                avg_lfct = mean(umca_slc, na.rm = T),
                tot_lfct = sum(umca_slc, na.rm = T))
summary(umca_in_plots)
# length(unique(umca_plots$plotid))# 183 plots with bay laurel
# umca_plots <- filter(plots_umca, plot != "BUSH01")
rm(plots_umca)
```

```{r calculate plot level leaf counts}
umca_plots <- left_join(
      umca_plots,
      plot_visit_dates,
      by = c("plotid","year")
)
summary(umca_plots)

umca_in_plots <- left_join(
      umca_in_plots,
      plot_visit_dates,
      by = c("plotid","year")
)
summary(umca_in_plots)
```

@@

# Join the environmental data to the bay laurel infection data
umca_plots <- left_join(umca_plots, plots_env, by = "plotid")
summary(umca_plots)
# filter(umca_plots, is.na(lon_utm))
# NAs due to no topo data for abandoned SUGAR02; remove SUGAR02
# umca_plots <- na.omit(umca_plots)
umca_plots <- rename(umca_plots, sample_year = year)
# umca_plots$plotid <- tolower(umca_plots$plotid)
umca_in_plots <- left_join(umca_in_plots, plots_env, by = "plotid")
umca_in_plots <- rename(umca_in_plots, sample_year = year)
# Join rainy season and summer temperatures to bay laurel infection data.
## I need to rename temperature variables to differentiate rainy vs. dry season
rs_temps <- as.tbl(rs_temps)
rs_temps$plotid <- tolower(rs_temps$plotid)
summary(rs_temps)
rs_temps <- rename(rs_temps, 
                   hrs_14_20_rs = hrs_14_20, hrs_abv25rs = hrs_abv25,
                   hrs_abv20_rs = hrs_abv20, hrs_blw10rs = hrs_blw10,
                   avg_tmax_rs = avg_tmax, avg_tmin_rs = avg_tmin)

@@

# Now I can join the data without duplicating variable names
umca_plots <- left_join(umca_plots, rs_temps, by = c("plotid", "sample_year"))
umca_plots <- left_join(umca_plots, summer_temps, by = c("plotid", "sample_year"))
summary(umca_plots)# 7 NAs introduced into the temperature variables
filter(umca_plots, is.na(avg_tmax_rs))# NAs due to records from 2003
## I don't think these will end up in my analysis, but no reason to remove them at this time.
umca_in_plots <- left_join(umca_in_plots, rs_temps, by = c("plotid", "sample_year"))
umca_in_plots <- left_join(umca_in_plots, summer_temps, by = c("plotid", "sample_year"))

@@

umca_in_plots$sample_year <- as.character(umca_in_plots$sample_year)
umca_in_plots <- left_join(umca_in_plots, rain_total, by = c("plotid", "sample_year"))
umca_in_plots <- left_join(umca_in_plots, rain_days, by = c("plotid", "sample_year"))

@@

umca_plots <- left_join(umca_plots, os.us.diversity, by = c("plotid"))
umca_in_plots <- left_join(umca_in_plots, os.us.diversity, by = c("plotid"))
summary(umca_plots)
unique(filter(umca_plots, is.na(H.2014))$plotid)
## NAs due to plots abandoned/lost between 2005 and 2014; 4 unique plots
filter(umca_plots, is.na(us.H.2005))
## NAs due to plot having '0' for understory richness
filter(umca_plots, is.na(us.H.2011))
## NAs due to 2 plots being abaondoned prior to 2011 & one plot having '0' for understory richness

@@

# Add bay laurel infection data to oak plots
names(umca_plots)
oak_sod <- left_join(oak_sod, 
                     select(umca_plots, plotid, sample_year, 
                            avg_lfct, tot_lfct ),
                     by = c("plotid","sample_year"))
names(umca_in_plots)
umca_in_plots$sample_year <- as.character(umca_in_plots$sample_year)
oak_sod$sample_year <- as.character(oak_sod$sample_year)
oak_sod <- left_join(oak_sod,
                     select(umca_in_plots, plotid, sample_year,
                            tot_bay, infected_bay_ct),
                     by = c("plotid","sample_year"))
summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
plot_visit_dates$year <- as.character(plot_visit_dates$year)
oak_sod <- left_join(
      oak_sod,
      plot_visit_dates,
      by = c("plotid","sample_year"="year"))

filter(oak_sod, is.na(tot_bay), sample_year != 2003)
# Change NAs of bay laurel in oak only plots to zeroes
# oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
oak_sod$infected_bay_ct[is.na(oak_sod$infected_bay_ct)] <- 0
oak_sod$tot_bay[is.na(oak_sod$tot_bay)] <- 0

@@

rs_temps
summer_temps
oak_sod$sample_year <- as.integer(oak_sod$sample_year)
oak_sod <- left_join(oak_sod, rs_temps, by = c("plotid", "sample_year"))
oak_sod <- left_join(oak_sod, summer_temps, by = c("plotid", "sample_year"))
oak_sod
names(oak_sod)
